---
title: "WUP_04"
author: "Kacey Bates"
date: "September 17, 2024"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
GSScsv <- read_csv("GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
basicCN <- ggplot(GSScsv,aes(x=DEGREE,y=WORDSUM))
```
```{r}
basicCN + geom_boxplot()
```


Based on the box plot, you can see that there is a difference shown, graduates have a higher Q1 than high schools Q3.Junior colleges Q1 is also just under LT high schools Q3.


```{r}
boxplot(WORDSUM~DEGREE,data=GSScsv)
```
### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSScsv)
```


Graduate has a Q1 of 10 compared to the Q3 of LT high school is 8.25 and the Q3 of high school is 9 even though it has a high number of 85 people. The number of people in high school is almost the combined number of all other degree's showing a difference between the values.

## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSScsv)
summary.aov(model1)
```


The p-value is 0.000236, meaning that there is a statistical difference within these results. Further testing is required to see the results.

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 17, 2, 2))
plot(TukeyHSD(model1, conf.level = 0.98), las=2)
```

The difference within this graph is clearly there, as the LT high school-graduate and the high school graduates are below 0 showing that there is a significant difference when it is compared to the other degree's.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

A relationship was found within this data, we visually saw a difference with the GRADUATE's being significantly higher than the high school DEGREE and the LT high school DEGREE. To fully see this we saw the numbers showing that there was a difference as the Q1 of the graduate was 10 while the Q3 of LT high school is 8.25 and the Q3 of HIGH SCHOOL is 9. This was proven with the one way anova showing a p-value of 0.000236 proving something was happening. After playing around with the code we were able to see a significant difference between two levels within this data, those levels being the high school and LThigh school as they are below 0.